---
title: "capstone_supernova_classification"
output: pdf_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Question 1:

How can we best what are the most impactful variables and best classification model to use when it comes to classiyfing supernovas with the given data using a decision/classification tree.



## Imports and data init
```{r}
library(tidyverse)
library(randomForest)
library(dplyr)
library(rpart)
library(rpart.plot)
library(party)

lightcurve_data_b_1 <- read.csv("Data/plasticc_test_set_batch1.csv")
lightcurve_train <- read.csv("Data/plasticc_train_lightcurves.csv")
meta_data_test <- read.csv("Data/plasticc_test_metadata.csv")
meta_data_train <- read.csv("Data/plasticc_train_metadata.csv")
```


## Helper function init
```{r pressure, echo=FALSE}
get_zero_point_flux_mag <- function(filter){
  f_0 = 0
  #m_0 = 0
    if(filter == "u"){
      f_0 = 1810#Jy
      #m_0 = 23.9#Jy
  } else if (filter == "g"){ #SDSS passbands calculated with Vega as zero point reference
    f_0 = 3640
    #m_0 = 25.1
  } else if (filter == "r"){
    f_0 = 3080
    #m_0 = 24.4
  } else if (filter == "i"){
    f_0 = 2550
    #m_0 = 24.1
  } else if (filter == "z"){
    f_0 = 1800
    #m_0 = 22.8
  } else if (filter == "y"){ 
    f_0 = 1080
    #m_0 = 20.9
  }
  else{
    f_0 = NULL
    #m_0 = NULL
  }
  return(f_0)
}

calculate_apparent_magnitude <- function(flux, filter){
  # We assume the flux of the objects are already converted to Jy
  f <- abs(flux) # to avoid NaNs from log(-x)
  return(-2.5 * log10(f/get_zero_point_flux_mag(filter)))
}

calculate_absolute_magnitude <- function(flux, distance, filter) {
  # We assume the flux of the objects are in nanomaggies given their size.
  # We assume distance is in the modulus scale
  # We assume filter is one of (u g r i z y)
  # Values of f_0 are using SDSS standards.
  f <- abs(flux) * 3.631 * 10^(-6) # to avoid NaNs from log(-x) and to convert to Jy from nanomaggies
  apparent_magnitude <- calculate_apparent_magnitude(f, filter)
  return(apparent_magnitude - distance)
}

calculate_luminosity <- function(flux, distmod){
  # We assume flux is in nanomaggies
  # We assume distmod is in distmod
  distance_parsec = 10^(distmod/5 + 1)
  flux_erg = flux * 3.631e-6 * 10^(-0.4*distmod) 
  return(4*pi * (distance_parsec)^2 * flux_erg)
}

get_lightcurve_summary <- function(lightcurve_data_param, potential_supernovas_param){
  
  # lightcurve_data_param <- light_train
  # potential_supernovas_param <- potential_supernovas_train
  
  filter_lightcurve_data <- lightcurve_data_param %>% 
    filter(flux != 0, object_id %in% potential_supernovas_param$object_id) 
  
  passbands = c("u", "g", "r", "i", "z", "y")
  
  mut_filter_lightcurve_data <- filter_lightcurve_data %>% mutate(passband_letter = passbands[passband + 1])
  
  mut_meta_lightcurve_merge <- merge(mut_filter_lightcurve_data, potential_supernovas_param, by = "object_id", all.x = TRUE)
  
  suppressWarnings({
  mut_meta_lightcurve_merge_abs_mag <- mut_meta_lightcurve_merge %>% rowwise() %>% mutate(
  absolute_magnitude = calculate_absolute_magnitude(flux, true_distmod, passband_letter),
  luminosity = calculate_luminosity(flux, true_distmod))
  })
  
  df_max_luminosity <- mut_meta_lightcurve_merge_abs_mag %>% group_by(object_id) %>% filter(luminosity == max(luminosity)) %>% summarise(object_id = object_id, max_luminosity = max(luminosity), mjd_max_luminosity = mjd)
  
  df_50pct_luminosity <- mut_meta_lightcurve_merge_abs_mag %>% group_by(object_id) %>% filter(luminosity <= 0.5 * max(luminosity), mjd > mut_meta_lightcurve_merge_abs_mag[match(max(luminosity),                                        mut_meta_lightcurve_merge_abs_mag$luminosity), ]$mjd) %>% summarize(
    object_id = object_id[1],
    val_50pct_luminosity = luminosity[1],
    mjd_50pct_luminosity = mjd[1]
    ) %>% select(object_id, mjd_50pct_luminosity, val_50pct_luminosity)
  
    merge_summary <- mut_meta_lightcurve_merge_abs_mag %>% group_by(object_id) %>% summarise(
    id = object_id,
    class_label = true_target,
    max_absolute_magnitude = max(absolute_magnitude),
    min_absolute_magnitude = min(absolute_magnitude),
    min_luminosity = min(luminosity),
    mjd_range = max(mjd) - min(mjd),
    vpec = true_vpec) %>% distinct(object_id, .keep_all = TRUE)
    
    
    merged_dfs <- merge(df_max_luminosity, df_50pct_luminosity, by = "object_id")
    merge_summary_complete <- merge(merge_summary, merged_dfs, by = "object_id") %>% mutate(days_to_50pct_luminosity = mjd_50pct_luminosity - mjd_max_luminosity)

  return(merge_summary_complete)
}

```


The provided data-set defines 14 class labels that correspond to differnt transient events. The ones were are interested in are:
  class 6: SNIa (Type Ia supernova)
  class 16: SNIbc (core-collapse supernova Type Ibc)
  class 42: SNIIn (Type IIn supernova)
  class 52: SNII (core-collapse supernova Type II)
  class 53: SNIIL (Type II-L supernova)
  class 62: SNIIb (Type IIb supernova)

## Object classification
```{r}
light_train <- lightcurve_train
light_test <- lightcurve_data_b_1


potential_supernovas_train <- meta_data_train %>% filter(true_target <= 95, true_target %in% c(6, 16, 42, 52, 53, 62)) # these values represent (SNIa, SNIbc, SNIIn, SNII, SNIIL, SNIIb)
## Research shows that only 14 class labels are recognized, anything greater than this is not a value that can be classified by this algorithm.

potential_supernovas_test <- meta_data_test %>% filter(true_target <= 95, true_target %in% c(6, 16, 42, 52, 53, 62))


# potential_supernovas_train %>% ggplot() + aes(x=true_target) + geom_bar() + ggtitle("Distribution of true_target for objects that are most likley supernovas (train)")
# 
# potential_supernovas_test %>% ggplot() + aes(x=true_target) + geom_bar() + ggtitle("Distribution of true_target for objects that are most likley 

#meta_data %>% ggplot() + aes(x=true_target) + geom_bar() + ggtitle("Distribution of true_target for objects in meta data")


train_summary <- get_lightcurve_summary(light_train, potential_supernovas_train)
test_summary <- get_lightcurve_summary(light_test, potential_supernovas_test)

classification_model_rpart = rpart(class_label ~ max_absolute_magnitude + min_absolute_magnitude + max_luminosity + min_luminosity + days_to_50pct_luminosity + vpec, data = train_summary, method = 'class')
classification_model_rforest <- randomForest(class_label ~ max_absolute_magnitude + min_absolute_magnitude + max_luminosity + min_luminosity + days_to_50pct_luminosity + vpec, data = train_summary)

classification_model_ctree <- ctree(class_label ~ max_absolute_magnitude + min_absolute_magnitude + max_luminosity + min_luminosity + days_to_50pct_luminosity + vpec, data = train_summary)
classification_model_cforest <- cforest(class_label ~ max_absolute_magnitude + min_absolute_magnitude + max_luminosity + min_luminosity + vpec, data = train_summary)

predict_test_rpart <- predict(classification_model_rpart, test_summary, type = "class")
predict_test_rforest <- predict(classification_model_rforest, test_summary, type = "class")
predict_test_ctree <- predict(classification_model_ctree, newdata = test_summary)
predict_test_cforest <- predict(classification_model_cforest, newdata = test_summary)


confusion_matrix_rpart <-  as.matrix(table(test_summary$class_label, predict_test_rpart))
confusion_matrix_rforest <- as.data.frame.matrix(table(test_summary$class_label, predict_test_rforest))
confusion_matrix_ctree <- as.matrix(table(test_summary$class_label, predict_test_ctree))
confusion_matrix_cforest <- as.data.frame.matrix(table(test_summary$class_label, predict_test_cforest))

percent_correct_rpart <- sum(diag(confusion_matrix_rpart)) / sum(confusion_matrix_rpart)
percent_correct_rforest <- sum(diag(as.matrix(confusion_matrix_rforest))) / sum(as.matrix(confusion_matrix_rforest))
percent_correct_ctree <- sum(diag(confusion_matrix_ctree)) / sum(confusion_matrix_ctree)
percent_correct_cforest <- sum(diag(as.matrix(confusion_matrix_cforest))) / sum(as.matrix(confusion_matrix_cforest))



final_estimations <- data.frame(x = c("rpart", "rforest", "ctree","cforest"), y = c(percent_correct_rpart, percent_correct_rforest, percent_correct_ctree, percent_correct_cforest))

ggplot(data = final_estimations, aes(x = x, y = y)) +
  geom_bar(stat = "identity") +
  labs(x = "Model", y = "Percent Sucess") +
  ggtitle("Graph of model accuracy")

```


